Approximate State Abstraction

### 2016-01-30 paper/paper.tex
Added proof for lemma 2 but it just seems like simulation lemma

Inspired by the model-irrelevance abstraction, we consider the approximate case. We let $\phi_{model}$ define a type of abstraction that, for fixed $\epsilon_\mathcal{R}$ and $\epsilon_\mathcal{T}$, satisfies:
\begin{multline}
\ep_{model}(s_1) = \ep_{model}(s_2) \rightarrow \\
\forall_a \left|\mathcal{R}(s_1, a) - \mathcal{R}(s_2, a)\right| \leq \epsilon_\mathcal{R}\ \wedge\ \\
\forall_{a,s'} \left|\mathcal{T}(s_1, a,s') - \mathcal{T}(s_2, a,s')\right| \leq \epsilon_\mathcal{T}
\end{multline}

@@

\begin{equation}
{M_G}' =  \langle \mathcal{S}_G, \mathcal{A}, \mathcal{R}_G', \mathcal{T}_G', \gamma \rangle
\end{equation}

Where
\begin{align}
{\mcR_G}'(s,a) = \mcR_A(\ep_m(s), a)\\
{\mcT_G}'(s,a, s') = \frac{\mcT_A(\ep_m(s), a, \ep_m(s'))}{|X(s')|}
\end{align}

where $X(s') = \{g \in \mathcal{S}_G | \ep_m(g) = \ep_m(s')\}$.

It is easily verifiable that ${\mcT_G}'$ induces a valid probability distribution.

@@

Similarly, the transitions of states in ${M_G}'$ that are aggregated in $M_A$ are identical:
\begin{align*}
\forall{s1_G, s2_G, {s_G}' \in \mcS_G, a}\\
\left[ \epM(s1_G) = \epM(s2_G) \rightarrow {\mcT_G}'(s1_G, a, {s_G}') = {\mcT_G}'(s2_G, a, {s_G}') \right]
\end{align*}

Now consider applying to $M_G$' the state aggregation scheme, $\phi_m$, which satisfies the following:
\begin{align*}
\forall{a, s_G \in \mcS_G}  [&{\mcR_G}'(s1_G, a) = {\mcR_G}'(s2_G, a) \wedge \\
&{\mcT_G}'(s1_G, a, {s_G}') = {\mcT_G}'(s2_G, a, {s_G}') ]  \rightarrow \\
\phi_m(s1_G) = \phi_m(s2_G)
\end{align*}

@@

 \cite{li2006towards} demonstrate that abstraction schemes of the $\pM$ type maintain Q values so we know that $\forall{s_G \in \mcS} Q_{G'}(s_G) = Q_{A}(s_G)$.

Knowing that the Q values in $M_A$ are identical to those in ${M_G}'$, we now demonstrate that those of $M_G$ are boundedly close to those of ${M_G}'$ and then conclude by applying Lemma 1 (which states similar Q values in the abstract and ground MDP results in near optimal behavior of the optimal abstract policy applied to the ground MDP).

Let the maximum upper bound between optimal behavior in the noisy and ground MDP be called $B$.

\begin{align*}
B_{s} &= \max_{s,a}|Q'(s,a)-Q(s,a)|\\
%B_{s} &= \max_{s,a}|R'(s,a) - R(s,a) + \gamma \sum_{s'}\left[T'(s,a,s')\max_{a'}Q'(s',a')-T(s,a,s')\max_{a'}Q(s', a')\right]|\\
&\forall s' \text{ WLOG } T'(s,a,s')\max_{a'}Q'(s',a') \geq T(s,a,s')\max_{a'}Q(s',a')\\
\end{align*}
Applying max is a non-expansion \enote{Probably want a citation and or explanation for this}:
\begin{multline}
=\max_{s,a}(|R'(s,a)-R(s,a)| + \gamma \sum_{s'} \\
\max_{a'} [(T(s,a,s')+T'(s,a,s')- \\ T(s,a,s'))Q'(s', a')-T(s,a,s')Q(s',a')])
\end{multline}
%&\leq \max_{s,a}|R'_s-R_s|+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')+\epsilon_{\mathcal{T}})Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}}+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')Q'(s',a')+\epsilon_{\mathcal{T}}Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}} + \max_{s,a} \gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')(Q'(s',a')-Q(s',a'))+\epsilon_{\mathcal{T}}Q'(s',a') \right]\\
\begin{align*}
&\leq \epsilon_{\mathcal{R}}+\gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{\text{max}} + \gamma \sum_{s'}\left[ T(s,a,s') B_s\right]\\
&\leq \epsilon_{\mathcal{R}} + \gamma B_{s} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{max}
\end{align*}

Applying $Q'_{max}=V'_{max} \leq V_{max}=\frac{1}{1-\gamma}$:
\begin{align*}
B_{s} &\leq 
\frac
{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}}
{(1-\gamma)^2}
\end{align*}

Therefore:
\begin{align*}
\forall_{s,a} : |Q'(s,a) - Q(s,a)| &\leq \max_{s,a}|Q'(s,a)-Q(s,a)| \\
&\leq  \frac{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}} }{(1-\gamma)^2}
\end{align*}

Consequently, from the below $\phi_Q^*$ results, we conclude that:
\begin{equation}
\forall_s : |V^{\pi^*_M}(s) - V^{\pi^*_{M'}}(s)| \leq \frac{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}}{(1-\gamma)^3}
\end{equation}

>>>>

Inspired by the model-irrelevance abstraction, we consider the approximate case. We let $\phi_{model}$ define a type of abstraction that, for fixed $\epsilon$, satisfies:
\begin{multline}
\ep_{model}(s_1) = \ep_{model}(s_2) \rightarrow \\
\forall_a \left|\mathcal{R}(s_1, a) - \mathcal{R}(s_2, a)\right| \leq \epsilon \wedge\ \\
\forall_{a,s'} \left|\mathcal{T}(s_1, a,s') - \mathcal{T}(s_2, a,s')\right| \leq \epsilon
\end{multline}

@@

\begin{equation}
{M_G}' =  \langle \mcS_G, \mathcal{A}, {\mcR_G}', {\mcT_G}', \gamma \rangle
\end{equation}

Where
\begin{align}
&{\mcR_G}'(s,a) = \mcR_A(\ep_m(s), a)\\
&{\mcT_G}'(s,a, s') = \frac{\mcT_A(\ep_m(s), a, \ep_m(s'))}{|X(s')|}
\end{align}

where $X(s') = \{g \in \mathcal{S}_G | \ep_m(g) = \ep_m(s')\}$.
\enote{Disambiguate this X from earlier X}

It is easily verifiable that ${\mcT_G}'$ is a probability distribution.

@@

The transitions of states in ${M_G}'$ that are aggregated in $M_A$ are identical :
\begin{align}
\forall{s1_G, s2_G, {s_G}' \in \mcS_G, a} \nonumber\\
\left[ \epM(s1_G) = \epM(s2_G) \rightarrow {\mcT_G}'(s1_G, a, {s_G}') = {\mcT_G}'(s2_G, a, {s_G}') \right] \wedge\label{eq:sameTrans}
\end{align}

Additionally, it is easily verified \enote{or not -- consider adding proof} that transitions and rewards in $M_G'$ are within $\epsilon$ of those in $M_G$:
\begin{align*}
\forall{s,s' \in \mcS_G, a} : |\mcT_G'(s,a,s') - \mcT_G(s,a,s')| \leq \epsilon\\
\forall{s \in \mcS_G, a} : |\mcR_G'(s,a) - \mcR_G(s,a)| \leq \epsilon
\end{align*}

\enote{If this is the case, why can't we just leverage the simulation lemma?...}

Now consider applying to $M_G$' the state aggregation scheme, $\phi_m$, which satisfies the following:
\begin{align}
\forall{a, s_G \in \mcS_G} \left[\epM(s1_G) =\epM(s2_G) \leftrightarrow \phi_m(s1_G) = \phi_m(s2_G) \right] \label{eq:sameR}
\end{align}

@@

 \cite{li2006towards} demonstrate that abstraction schemes of the $\pM$ type maintain Q values so we know that $\forall{s_G \in \mcS_G}  \left[ Q_{G'}(s_G) = Q_{A}(\epM(s_G)) \right]$ where $Q_{G'}$ is the Q function in ${M_G}'$ and $Q_G$

Knowing that the Q values in $M_A$ are identical to those in ${M_G}'$, we now demonstrate that those of $M_G$ are boundedly close to those of ${M_G}'$.

For the remainder of the proof, $s, s' \in \mcS_G$.

Let the maximum Q difference in ${M_G}'$ and $M_G$ be $B$:
\begin{align*}
B = &\max_{s,a}|Q_{G'}(s,a)-Q_{G}(s,a)|\\
%B_{s} &= \max_{s,a}|R'(s,a) - R(s,a) + \gamma \sum_{s'}\left[T'(s,a,s')\max_{a'}Q'(s',a')-T(s,a,s')\max_{a'}Q(s', a')\right]|\\
&\text{Applying the definition of Q values:}\\
%---Write out using Q definitions
B \leq &\max_{s,a} |\mcR_G'(s,a)-\mcR(s,a) +  \gamma \sum_{s'} \\
&\left[ \mcT_G'(s, a, s')\max_{a'}Q_G'(s', a') - \mcT_G(s, a, s')\max_{a'}Q_G(s', a')\right] |\\
%---Apply WLOG
& \text{ WLOG }T'(s,a,s')\max_{a'}Q'(s',a') \geq T(s,a,s')\max_{a'}Q(s',a')\\
\leq &\max_{s,a}(|\mcR_G'(s,a)-\mcR(s,a)| + \gamma \sum_{s'} \\
&\max_{a'} [(\mcT_G'(s,a,s'))Q_G'(s', a')-\mcT_G(s,a,s')Q_G(s',a')])\\
%---Apply epsilon to reward and transitions
\leq & \epsilon +\\
& \gamma \sum_{s'}  \max_{a'} [(\epsilon+ \mcT_G(s,a,s'))Q_G'(s', a')-\mcT_G(s,a,s')Q_G(s',a')])\\
%---Do algebra
= & \epsilon +\\
&\gamma \sum_{s'}  \max_{a'} [\epsilon Q_G'(s', a')+ \mcT_G(s,a,s')(Q_G'(s', a') - Q_G(s', a'))])
%---Pull terms out of some
&\leq \epsilon+\gamma |\mcS|\epsilon Q'_{\text{max}} + \gamma \sum_{s'}\left[ \mcT(s,a,s') B\right]\\
&\leq \epsilon + \gamma B + \gamma |\mcS|\epsilon Q'_{max}
\end{align*}

Applying $Q'_{max}=V'_{max} \leq V_{max}=\frac{1}{1-\gamma}$:
\begin{align*}
B &\leq 
\frac
{ \gamma |\mathcal{S} + 1 |\epsilon}
{(1-\gamma)^2}
\end{align*}

Therefore:
\begin{align*}
\forall_{s,a} : &|Q_A(s,a) - Q_G(s,a)| = |Q_G'(s,a) - Q_G(s,a)| \\
&\leq \max_{s,a}|Q_G'(s,a)-Q_G(s,a)| \\
&\leq  \frac{\gamma |\mathcal{S} + 1|\epsilon}{(1-\gamma)^2}
\end{align*}

Consequently, by the logic of the proof for Lemma 1, we conclude that:
\begin{equation}
\forall_s : |V^{\pi^*_M}(s) - V^{\pi^*_{A}}(s)| \leq
\frac{2\gamma|\mcS + 1|\epsilon}{(1-\gamma)^3}
\end{equation}
